---
title: "Bedinvetmab Market Analysis and Adverse Event Rates"
author: ""
date: "`r Sys.Date()`"
output: 
  word_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)



rm(list = ls())
set.seed(2612)

# Set working directory to project root
here::i_am("rate_simulation.Rmd")

# Load packages (install if needed)
required_packages <- c("tidyverse", "readxl", "here", "survival", "truncnorm", 
                       "forcats", "viridis", "tictoc", "janitor", "scales", 
                       "progressr", "gt", "jsonlite", "lubridate")

install_if_missing <- function(packages) {
  missing <- packages[!packages %in% rownames(installed.packages())]
  if (length(missing) > 0) {
    install.packages(missing, dependencies = TRUE)
  }
}

install_if_missing(required_packages)
invisible(lapply(required_packages, library, character.only = TRUE))
```

## Load and Filter AE Data

Starting with adverse event data for bedinvetmab through June 2025. This date was chosen as it is the most recent update for the number of Bedinvetmab doses sold (>30 million).

```{r load-data, include = FALSE}
ae_df <- read_rds(here("data", "processed", "matched_data.rds")) %>%
  filter(drug == "bedinvetmab",
         date <= as.Date("2025-06-30"))
```

## Country-Specific Reporting Rates

Using the Montiero pharmacovigilence study (2025) as a basis, they make it clear that different countries have different spontaneous reporting rates for bedinvetmab. Though they do not provide explicit units sold on a per country basis, using the spontaneous rates of reporting and the numbers of ADRs available from the EudraVigilance, one can make an informed prediction. EU countries were given the average of the known EU countries. Non-EU countries were called "Rest of World" and given the average across all countries.

```{r reporting-rates, include = FALSE}
reporting_rates <- tribble(
  ~country_group, ~ae_per_10k_doses,
  "United States", 18.18,
  "United Kingdom", 14.61,
  "Germany", 6.06,
  "Spain", 5.41,
  "France", 3.22,
  "Italy", 3.22,
  "Canada", 18.72,
  "Australia", 13.86,
  "EU_combined", mean(c(3.22, 6.06, 5.41, 3.22)),
  "Rest_of_World", 9.48
)
```

## Life Expectancy by Size and Age

The Monteiro study also provided weight and and age data of the dogs reporting ADR.  Using Montoya et al. 2019 data (Supplementary Table 8) for size-specific life expectancy estimates, survival curves for the simulated population.

```{r life-tables, include = FALSE}
life_table_by_size <- tribble(
  ~size_group, ~age_lower, ~age_upper, ~e_x_years, ~ci_lower, ~ci_upper,
  # Toy breeds
  "Toy", 0, 1, 13.58, 13.52, 13.63,
  "Toy", 1, 2, 12.97, 12.92, 13.02,
  "Toy", 2, 3, 12.09, 12.04, 12.14,
  "Toy", 3, 4, 11.19, 11.14, 11.24,
  "Toy", 4, 5, 10.28, 10.24, 10.33,
  "Toy", 5, 6, 9.38, 9.34, 9.43,
  "Toy", 6, 7, 8.49, 8.45, 8.54,
  "Toy", 7, 8, 7.62, 7.58, 7.67,
  "Toy", 8, 9, 6.77, 6.73, 6.81,
  "Toy", 9, 10, 5.97, 5.92, 6.01,
  "Toy", 10, 11, 5.22, 5.17, 5.26,
  "Toy", 11, 12, 4.52, 4.48, 4.57,
  "Toy", 12, 13, 3.88, 3.84, 3.93,
  "Toy", 13, 14, 3.32, 3.28, 3.37,
  "Toy", 14, 15, 2.85, 2.80, 2.91,
  "Toy", 15, 16, 2.47, 2.41, 2.53,
  "Toy", 16, 17, 2.23, 2.15, 2.31,
  "Toy", 17, 99, 2.08, 1.97, 2.18,
  # Small breeds
  "Small", 0, 1, 13.79, 13.74, 13.83,
  "Small", 1, 2, 13.02, 12.98, 13.06,
  "Small", 2, 3, 12.10, 12.07, 12.14,
  "Small", 3, 4, 11.18, 11.14, 11.21,
  "Small", 4, 5, 10.26, 10.22, 10.30,
  "Small", 5, 6, 9.35, 9.31, 9.39,
  "Small", 6, 7, 8.44, 8.41, 8.48,
  "Small", 7, 8, 7.56, 7.52, 7.59,
  "Small", 8, 9, 6.69, 6.65, 6.72,
  "Small", 9, 10, 5.88, 5.85, 5.91,
  "Small", 10, 11, 5.11, 5.08, 5.15,
  "Small", 11, 12, 4.41, 4.37, 4.44,
  "Small", 12, 13, 3.76, 3.73, 3.79,
  "Small", 13, 14, 3.20, 3.16, 3.23,
  "Small", 14, 15, 2.73, 2.69, 2.77,
  "Small", 15, 16, 2.36, 2.31, 2.40,
  "Small", 16, 17, 2.10, 2.05, 2.16,
  "Small", 17, 99, 1.97, 1.89, 2.04,
  # Medium breeds
  "Medium", 0, 1, 12.94, 12.88, 12.99,
  "Medium", 1, 2, 12.14, 12.09, 12.20,
  "Medium", 2, 3, 11.23, 11.18, 11.28,
  "Medium", 3, 4, 10.31, 10.26, 10.36,
  "Medium", 4, 5, 9.38, 9.33, 9.44,
  "Medium", 5, 6, 8.49, 8.44, 8.54,
  "Medium", 6, 7, 7.58, 7.53, 7.63,
  "Medium", 7, 8, 6.71, 6.66, 6.76,
  "Medium", 8, 9, 5.87, 5.82, 5.91,
  "Medium", 9, 10, 5.09, 5.05, 5.14,
  "Medium", 10, 11, 4.36, 4.32, 4.41,
  "Medium", 11, 12, 3.73, 3.69, 3.77,
  "Medium", 12, 13, 3.16, 3.11, 3.20,
  "Medium", 13, 14, 2.69, 2.64, 2.73,
  "Medium", 14, 15, 2.32, 2.26, 2.38,
  "Medium", 15, 16, 2.06, 1.99, 2.13,
  "Medium", 16, 17, 1.90, 1.80, 1.99,
  "Medium", 17, 99, 1.90, 1.75, 2.05,
  # Large breeds
  "Large", 0, 1, 11.70, 11.66, 11.73,
  "Large", 1, 2, 10.97, 10.93, 11.00,
  "Large", 2, 3, 10.08, 10.05, 10.12,
  "Large", 3, 4, 9.19, 9.15, 9.22,
  "Large", 4, 5, 8.28, 8.25, 8.31,
  "Large", 5, 6, 7.40, 7.37, 7.43,
  "Large", 6, 7, 6.54, 6.51, 6.57,
  "Large", 7, 8, 5.71, 5.68, 5.74,
  "Large", 8, 9, 4.93, 4.90, 4.96,
  "Large", 9, 10, 4.22, 4.19, 4.25,
  "Large", 10, 11, 3.62, 3.59, 3.65,
  "Large", 11, 12, 3.08, 3.05, 3.11,
  "Large", 12, 13, 2.62, 2.58, 2.65,
  "Large", 13, 14, 2.25, 2.21, 2.28,
  "Large", 14, 15, 1.98, 1.94, 2.02,
  "Large", 15, 16, 1.79, 1.73, 1.85,
  "Large", 16, 17, 1.66, 1.58, 1.75,
  "Large", 17, 99, 1.57, 1.44, 1.69,
  # Giant breeds
  "Giant", 0, 1, 9.70, 9.54, 9.85,
  "Giant", 1, 2, 8.93, 8.78, 9.08,
  "Giant", 2, 3, 8.08, 7.93, 8.23,
  "Giant", 3, 4, 7.21, 7.06, 7.36,
  "Giant", 4, 5, 6.37, 6.22, 6.51,
  "Giant", 5, 6, 5.59, 5.44, 5.73,
  "Giant", 6, 7, 4.87, 4.72, 5.01,
  "Giant", 7, 8, 4.21, 4.07, 4.36,
  "Giant", 8, 9, 3.64, 3.50, 3.79,
  "Giant", 9, 10, 3.18, 3.02, 3.34,
  "Giant", 10, 11, 2.82, 2.64, 2.99,
  "Giant", 11, 12, 2.52, 2.31, 2.72,
  "Giant", 12, 13, 2.32, 2.06, 2.57,
  "Giant", 13, 99, 2.08, 1.75, 2.41
) %>%
  mutate(
    sd_years = (ci_upper - ci_lower) / (2 * 1.96),
    e_x_months = e_x_years * 12,
    sd_months = sd_years * 12
  )
```

## Market Composition by Country

Making the assumption that there is a fixed correlation between ADR reports and market size; and knowing the country-specific reporting rates, we can back-calculate the market composition. This provides an estimate for where most doses are being used.

```{r country-analysis}
eu_countries <- c(
  "Austria", "Belgium", "Bulgaria", "Croatia", "Czech Republic",
  "Denmark", "Estonia", "Finland", "France", "Germany", "Greece",
  "Hungary", "Iceland", "Ireland", "Italy", "Lithuania", "Luxembourg",
  "Netherlands", "Norway", "Poland", "Portugal", "Slovakia", "Slovenia",
  "Spain", "Sweden", "Switzerland", "Andorra"
)

analysis_data <- ae_df %>%
  group_by(dog_id) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(country) %>%
  summarise(n_aes = n(), .groups = "drop") %>%
  mutate(
    country_group = case_when(
      country == "United States" ~ "United States",
      country == "United Kingdom" ~ "United Kingdom", 
      country == "Canada" ~ "Canada",
      country == "France" ~ "France",
      country == "Germany" ~ "Germany",
      country == "Italy" ~ "Italy",
      country == "Spain" ~ "Spain",
      country == "Australia" ~ "Australia",
      country %in% eu_countries ~ "EU_combined",
      TRUE ~ "Rest_of_World"
    )
  ) %>%
  group_by(country_group) %>%
  summarise(total_aes = sum(n_aes), .groups = "drop") %>%
  left_join(reporting_rates, by = "country_group") %>%
  mutate(
    license_date = case_when(
      country_group == "United States" ~ as_date("2023-05-05"),
      country_group == "United Kingdom" ~ as_date("2020-11-10"),
      country_group == "Canada" ~ as_date("2023-03-08"),
      country_group == "Australia" ~ as_date("2022-09-01"),
      country_group %in% c("France", "Germany", "Italy", "Spain", "EU_combined") ~ 
        as_date("2020-11-10"),
      country_group == "Rest_of_World" ~ as_date("2022-01-01")
    ),
    estimated_doses = total_aes / (ae_per_10k_doses / 10000),
    market_share = estimated_doses / sum(estimated_doses)
  ) %>%
  arrange(desc(market_share))

print(analysis_data)
```

## Uptake Trends

It's important to model the uptake of the drug following licensing (since there is a big variation in licensing dates around the world). 

```{r empirical-uptake}
empirical_uptake <- ae_df %>%
  mutate(
    year_month = floor_date(date, "month"),
    country_group = case_when(
      country == "United States" ~ "United States",
      country == "United Kingdom" ~ "United Kingdom",
      country == "Canada" ~ "Canada",
      country == "France" ~ "France",
      country == "Germany" ~ "Germany",
      country == "Italy" ~ "Italy",
      country == "Spain" ~ "Spain",
      country == "Australia" ~ "Australia",
      country %in% eu_countries ~ "EU_combined",
      TRUE ~ "Rest_of_World"
    )
  ) %>%
  count(country_group, year_month, name = "n_aes") %>%
  left_join(analysis_data %>% select(country_group, license_date), by = "country_group") %>%
  filter(year_month >= license_date) %>%
  left_join(reporting_rates %>% select(country_group, ae_per_10k_doses), by = "country_group") %>%
  mutate(
    months_since_launch = as.numeric(interval(license_date, year_month) / months(1)),
    estimated_doses = n_aes / (ae_per_10k_doses / 10000)
  ) %>%
  group_by(country_group) %>%
  arrange(year_month) %>%
  mutate(cumulative_doses = cumsum(estimated_doses)) %>%
  ungroup()


```

 ## Logistic Growth Model

 Fitting logistic curves to model market uptake over time. We fix K (carrying
 capacity) based on expected saturation at 72 months and only fit the growth
 rate (r) and inflection point (t0).

```{r logistic-model}

# Calculate K assuming 72-month saturation
calculate_K <- function(current_cumulative, months_available) {
  t_saturation <- 72
  r_assumed <- 0.1
  t0_assumed <- t_saturation / 2
  
  # Where should we be on S-curve now?
  expected_prop <- 1 / (1 + exp(-r_assumed * (months_available - t0_assumed)))
  
  # K = current / expected_proportion
  K_estimated <- current_cumulative / expected_prop
  
  return(K_estimated)
}

# Fit only r and t0 with K fixed
fit_logistic_fixed_K <- function(data, K_fixed) {
  max_time <- max(data$months_since_launch)
  
  tryCatch({
    nls(
      cumulative_doses ~ K_fixed / (1 + exp(-r * (months_since_launch - t0))),
      data = data,
      start = list(r = 0.1, t0 = max_time * 0.5),
      control = nls.control(maxiter = 200)
    )
  }, error = function(e) NULL)
}

# Apply fixed-K approach
logistic_params <- empirical_uptake %>%
  group_by(country_group) %>%
  summarise(
    months_available = max(months_since_launch),
    current = max(cumulative_doses),
    K = calculate_K(current, months_available),
    model = list(fit_logistic_fixed_K(cur_data(), K)),
    r = ifelse(!is.null(model[[1]]), coef(model[[1]])["r"], 0.1),
    t0 = ifelse(!is.null(model[[1]]), coef(model[[1]])["t0"], months_available * 0.5),
    pct_capacity = 100 * current / K,
    .groups = "drop"
  )

# Merge logistic params back into analysis data
analysis_data <- analysis_data %>%
  left_join(logistic_params, by = "country_group")

print("Fixed-K logistic parameters:")
print(logistic_params %>%
        select(country_group, months_available, K, r, t0, pct_capacity) %>%
        mutate(
          K = comma(round(K)),
          r = round(r, 3),
          t0 = round(t0, 1),
          pct_capacity = sprintf("%.1f%%", pct_capacity)
        ))

# Model fit quality
validation <- empirical_uptake %>%
  left_join(logistic_params %>% select(country_group, K, r, t0), by = "country_group") %>%
  mutate(predicted = K / (1 + exp(-r * (months_since_launch - t0)))) %>%
  group_by(country_group) %>%
  summarise(r_squared = cor(cumulative_doses, predicted, use = "complete.obs")^2, .groups = "drop")

print("\nR² with fixed K:")
print(validation %>% mutate(r_squared = round(r_squared, 3)))

```
### Visualize Model Fit

```{logistic-plot, fig.width=10, fig.height=6}

fitted_predictions_k <- empirical_uptake %>%
  left_join(logistic_params %>% select(country_group, K, r, t0), by = "country_group") %>%
  filter(!is.na(K)) %>%
  mutate(
    fitted_doses = K / (1 + exp(-r * (months_since_launch - t0)))
  )

comparison_data_k <- fitted_predictions_k %>%
  select(country_group, year_month, months_since_launch, 
         empirical = cumulative_doses, fitted = fitted_doses) %>%
  pivot_longer(cols = c(empirical, fitted), names_to = "source", values_to = "cumulative")

comparison_data_k %>%
  filter(country_group %in% c("United States", "United Kingdom", "EU_combined", "Canada")) %>%
  ggplot(aes(x = year_month, y = cumulative, color = source)) +
  geom_point(data = . %>% filter(source == "empirical"), size = 1.5, alpha = 0.7) +
  geom_line(data = . %>% filter(source == "fitted"), linewidth = 1.2) +
  facet_wrap(~country_group, scales = "free_y", ncol = 2) +
  scale_y_continuous(labels = comma) +
  scale_color_manual(
    values = c("empirical" = "#E74C3C", "fitted" = "#3498DB"),
    labels = c("Empirical (dose-corrected)", "Logistic Fit (Fixed K)")
  ) +
  labs(
    title = "Logistic Model: Empirical vs Fitted",
    subtitle = "Empirical data corrected for reporting rates",
    x = "Date",
    y = "Cumulative Dose-Equivalents",
    color = ""
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "top")
```

## Simulation Setup

Now we'll simulate a dog population that matches our known sales figures. Key parameters:

```{r simulation-params}
# Target to match
known_global_doses_sold <- 21000000
cutoff_date <- as.Date("2025-06-30")
dose_buffer_pct <- 0.10
dropout_monthly_yr1 <- 0.03

# Weight distribution (from prior analysis)
weight_mean <- 26.06
weight_sd <- 13.78

# Size categories based on weight
classify_size <- function(weight_kg) {
  case_when(
    weight_kg < 5.5 ~ "Toy",
    weight_kg < 11 ~ "Small",
    weight_kg < 26 ~ "Medium",
    weight_kg < 45 ~ "Large",
    TRUE ~ "Giant"
  )
}
```

## Generate Dog Population

We'll iteratively generate dogs until we hit our target dose count. Each dog gets assigned:
- Entry date (based on country market share and license dates)
- Weight (from truncated normal distribution)
- Age at start (uniform 1-14 years)
- Expected survival (from life tables)
- Treatment duration (with dropout)

```{r simulate-population, cache=TRUE}
dogs_list <- list()
iteration <- 1
total_doses_so_far <- 0
target_doses_consumed <- known_global_doses_sold * (1 - dose_buffer_pct)

while (total_doses_so_far < target_doses_consumed) {
  
  # Generate batch of dogs
  batch_size <- ifelse(iteration == 1, 50000, 10000)
  
  country_probs <- analysis_data$market_share
  names(country_probs) <- analysis_data$country_group
  
  batch <- tibble(
    dog_id = paste0("dog_", (iteration - 1) * batch_size + 1:batch_size),
    country_group = sample(names(country_probs), batch_size, 
                          replace = TRUE, prob = country_probs),
    weight_kg = rtruncnorm(batch_size, a = 1, b = 100, 
                          mean = weight_mean, sd = weight_sd),
    size_group = classify_size(weight_kg),
    age_at_start = runif(batch_size, 1, 14)
  ) %>%
    left_join(analysis_data %>% select(country_group, license_date), 
             by = "country_group")
  
  # Sample entry dates weighted toward early periods
  batch <- batch %>%
    rowwise() %>%
    mutate(
      days_since_launch = as.numeric(cutoff_date - license_date),
      # Beta distribution for entry timing
      entry_quantile = rbeta(1, 1.5, 3),
      days_after_launch = round(entry_quantile * days_since_launch),
      entry_date = license_date + days(days_after_launch)
    ) %>%
    ungroup() %>%
    select(-days_since_launch, -entry_quantile, -days_after_launch)
  
  # Lookup life expectancy
  batch <- batch %>%
    mutate(age_group_lower = floor(age_at_start)) %>%
    left_join(
      life_table_by_size %>% 
        select(size_group, age_lower, e_x_months, sd_months),
      by = c("size_group", "age_group_lower" = "age_lower")
    ) %>%
    mutate(
      survival_months = rtruncnorm(n(), a = 1, b = Inf, 
                                  mean = e_x_months, sd = sd_months),
      death_date = entry_date + months(round(survival_months))
    )
  
  # Calculate treatment duration with dropout
  batch <- batch %>%
    rowwise() %>%
    mutate(
      months_available = as.numeric(
        interval(entry_date, min(death_date, cutoff_date)) / months(1)
      ),
      # Dropout in first year only
      first_year_months = min(months_available, 12),
      dropout_months = sum(runif(first_year_months) < dropout_monthly_yr1),
      treatment_months = months_available - dropout_months,
      treatment_months = max(0, treatment_months),
      doses_by_cutoff = treatment_months
    ) %>%
    ungroup()
  
  dogs_list[[iteration]] <- batch
  total_doses_so_far <- sum(map_dbl(dogs_list, ~sum(.x$doses_by_cutoff)))
  
  iteration <- iteration + 1
  if (iteration > 100) break  # safety valve
}

dogs_final <- bind_rows(dogs_list)

# Trim to exact target
dogs_final <- dogs_final %>%
  arrange(entry_date) %>%
  mutate(cumulative_doses = cumsum(doses_by_cutoff)) %>%
  filter(cumulative_doses <= target_doses_consumed) %>%
  select(-cumulative_doses)
```

## Results by Country

```{r country-summary}
country_summary <- dogs_final %>%
  group_by(country_group) %>%
  summarise(
    n_dogs = n(),
    total_doses_consumed = sum(doses_by_cutoff),
    mean_age = mean(age_at_start),
    mean_weight = mean(weight_kg),
    median_weight = median(weight_kg),
    mean_duration = mean(doses_by_cutoff),
    median_duration = median(doses_by_cutoff),
    .groups = "drop"
  ) %>%
  left_join(analysis_data %>% select(country_group, total_aes), 
           by = "country_group") %>%
  mutate(
    total_doses_with_buffer = total_doses_consumed / (1 - dose_buffer_pct),
    ae_per_1000_dogs = total_aes / n_dogs * 1000
  ) %>%
  arrange(desc(n_dogs))

country_summary %>%
  select(country_group, n_dogs, mean_weight, mean_age, mean_duration, 
         total_aes, ae_per_1000_dogs) %>%
  mutate(
    n_dogs = comma(n_dogs),
    across(c(mean_weight, mean_age, mean_duration, ae_per_1000_dogs), ~round(., 1))
  ) %>%
  knitr::kable(caption = "Summary by Country/Region")
```

## Results by Size Group

Interesting to see how the different size categories break down.

```{r size-summary}
size_summary <- dogs_final %>%
  group_by(size_group) %>%
  summarise(
    n_dogs = n(),
    pct_of_total = 100 * n() / nrow(dogs_final),
    mean_weight = mean(weight_kg),
    mean_age = mean(age_at_start),
    mean_survival = mean(survival_months / 12),
    mean_treatment = mean(treatment_months),
    mean_doses = mean(doses_by_cutoff),
    .groups = "drop"
  ) %>%
  arrange(factor(size_group, levels = c("Toy", "Small", "Medium", "Large", "Giant")))

size_summary %>%
  mutate(across(where(is.numeric), ~round(., 1))) %>%
  mutate(n_dogs = comma(n_dogs)) %>%
  knitr::kable(caption = "Summary by Size Group")
```

## Global Summary

```{r global-summary}
total_dogs <- nrow(dogs_final)
total_doses <- sum(dogs_final$doses_by_cutoff)
total_doses_with_buffer <- total_doses / (1 - dose_buffer_pct)
total_aes <- sum(country_summary$total_aes, na.rm = TRUE)

cat(sprintf("
GLOBAL SUMMARY
==============

Dogs simulated: %s
Total doses consumed: %s
With 10%% buffer: %s
Target (known sales): %s
Accuracy: %.3f%%

Population characteristics:
  Mean weight: %.1f kg (median: %.1f kg)
  Mean age at start: %.1f years
  Mean treatment duration: %.1f months

Size distribution:
  Toy: %.1f%%
  Small: %.1f%%
  Medium: %.1f%%
  Large: %.1f%%
  Giant: %.1f%%

Adverse events:
  Total reports: %s
  Rate per 1,000 dogs: %.1f
  Rate per 10,000 doses: %.1f
",
  comma(total_dogs),
  comma(round(total_doses)),
  comma(round(total_doses_with_buffer)),
  comma(known_global_doses_sold),
  100 * abs(total_doses_with_buffer - known_global_doses_sold) / known_global_doses_sold,
  mean(dogs_final$weight_kg),
  median(dogs_final$weight_kg),
  mean(dogs_final$age_at_start),
  mean(dogs_final$doses_by_cutoff),
  100 * mean(dogs_final$size_group == "Toy"),
  100 * mean(dogs_final$size_group == "Small"),
  100 * mean(dogs_final$size_group == "Medium"),
  100 * mean(dogs_final$size_group == "Large"),
  100 * mean(dogs_final$size_group == "Giant"),
  comma(total_aes),
  total_aes / total_dogs * 1000,
  total_aes / total_doses_with_buffer * 10000
))
```

## ADR Rates by Country

Now let's calculate adverse drug reaction rates by preferred term (PT) across countries.

```{r adr-rates}
exposure_summary <- dogs_final %>%
  count(country_group, name = "n_exposed")

adr_summary <- ae_df %>%
  mutate(country_group = case_when(
    country == "United States" ~ "United States",
    country == "United Kingdom" ~ "United Kingdom",
    country == "Canada" ~ "Canada",
    country == "France" ~ "France",
    country == "Germany" ~ "Germany",
    country == "Italy" ~ "Italy",
    country == "Spain" ~ "Spain",
    country == "Australia" ~ "Australia",
    country %in% eu_countries ~ "EU_combined",
    TRUE ~ "Rest_of_World"
  )) %>%
  count(country_group, pt, name = "n_adrs")

adr_rates <- adr_summary %>%
  left_join(exposure_summary, by = "country_group") %>%
  mutate(rate_per_10000 = 10000 * n_adrs / n_exposed)

# Top 20 ADRs globally (excluding lack of efficacy)
top_pts <- adr_rates %>%
  filter(pt != "Lack of efficacy") %>%
  group_by(pt) %>%
  summarise(total_adrs = sum(n_adrs, na.rm = TRUE), .groups = "drop") %>%
  slice_max(total_adrs, n = 20) %>%
  pull(pt)

adr_top10 <- adr_rates %>%
  filter(pt %in% top_pts, pt != "Lack of efficacy")
```

## Heatmap of Top ADRs

```{r heatmap, fig.width=10, fig.height=8}
adr_plot <- adr_top10 %>%
  mutate(
    pt = str_to_title(pt) |> str_replace_all("_", " "),
    country_group = str_replace_all(country_group, "_", " ") |> str_to_title(),
    pt = fct_reorder(pt, -n_adrs),
    country_group = fct_reorder(country_group, n_exposed),
    text_color = ifelse(rate_per_10000 > 15, "white", "black")
  )

ggplot(adr_plot, aes(x = country_group, y = pt, fill = rate_per_10000)) +
  geom_tile(color = "grey85", linewidth = 0.4) +
  geom_text(
    aes(label = sprintf("%.1f", rate_per_10000), color = text_color),
    size = 3,
    fontface = "bold"
  ) +
  scale_color_identity() +
  scale_fill_viridis_c(
    option = "plasma",
    direction = -1,
    name = "ADR rate\n(per 10,000 dogs)",
    labels = label_number(accuracy = 1)
  ) +
  scale_x_discrete(position = "top", name = "Country or Region") +
  scale_y_discrete(name = "Adverse Reaction (Preferred Term)") +
  labs(
    title = "Top 10 Adverse Drug Reaction Rates by Country or Region",
    subtitle = "Rates calculated using simulated exposure denominators (per 10,000 dogs)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(face = "bold", size = 11, vjust = 0.5),
    axis.text.y = element_text(face = "bold", size = 11),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.key.width = unit(2.5, "cm"),
    legend.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold", size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 11, hjust = 0.5),
    plot.margin = margin(10, 20, 10, 10)
  )
```

## Notes

Key assumptions in this analysis:

- Weight distribution: truncated normal (mean=26.06kg, sd=13.78kg, range 1-100kg)
- Life expectancy: size-specific curves from Montoya et al. 2024
- Treatment dropout: 3% monthly in first year, 0% thereafter
- Dose buffer: 10% (accounting for wastage, stockpiling, etc.)
- Cutoff date: June 30, 2025

The simulation successfully matches known global sales within `r round(100 * abs(total_doses_with_buffer - known_global_doses_sold) / known_global_doses_sold, 3)`%.

```{r save-results, eval=FALSE}
# Uncomment to save results
# write_rds(dogs_final, here("output", "simulated_dogs_with_size.rds"))
# write_csv(size_summary, here("output", "size_summary.csv"))
# write_csv(country_summary, here("output", "country_summary.csv"))
```
